Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-10-11 manuscript/manuscript.Rmd
proofread manuscript (and auto-formatted R code)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(fig.pos = 'th') # Places figures at top or here
knitr::opts_chunk$set(out.width = '100%', dpi=300,
                      fig.width=8, fig.width=8) # Figure resolution and size
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment

options(knitr.table.format="latex") # For kable tables to work without setting format option

knitr::opts_chunk$set(echo=F, warning=F, message=F)#dev = "png", dev.args = list(type = "cairo-png")
 library("papaja")
library("bookdown")
library("rticles")
 library(here)
 library(tidyverse)
 library(brms)
library(tidybayes)
library(kableExtra)
library(viridis)
library(cowplot)
library(ggthemes)
library(reshape2)
library(knitr)
library(kableExtra)
library(metafor)
#r_refs("r-references.bib", append=F)
theme_set(theme_bw())
options(knitr.table.format = "pdf")

model_location="code/models"

```

```{r}

library(tidyverse)
library(testthat)
all_projects <- read_csv("https://raw.githubusercontent.com/vboyce/251-254-MA/main/data/raw_data.csv")

filtered <- all_projects |>
  mutate(sub_rep=ifelse(replicated_instructor_code==replicated_report_code, 
                 replicated_report_code, 
                 adjudicated_replication_code)) |> 
  filter(sub_rep %in% c(0,.25, .5)) |> 
  filter(include!="no") |> 
  filter(target_N <=200) |> 
  filter(academic_year %in% c("2015-2016", "2016-2017", "2017-2018", "2018-2019", 
                              "2019-2020", "2020-2021","2021-2022", "2022-2023")) 
# what we really want is whether there was a github, but that column is private and so not in the public raw data
# but 2015 is the first year where we have github links

#test_that("number of rows in filtered", {expect_equal(nrow(filtered),49)})

#given this sample, we then sought original replicator permission, 
#since it would involve sharing their repo and write-up with a new student
#which wouldn't have been anticipated at original classtime

#we heard back from 29 students of the 49 we attempted to contact 
#(1 we couldn't contact due to email bouncing) 
#2 responses were negative, rest were positive, for 27 options

approved <- c("krauss2003","yeshurun2003", "daffner2000", "ngo2019",
              "child2018", "schechtman2010", "payne2008", "paxton2012",
              "hart2018", "lewis2015", "tarampi2016", "jara-ettinger2022",
              "porter_2016_1", "hopkins2016", "birch2007_2", "gong2019",
              "correll2007", "pilditch2019", "daw2011", "dehaene2009",
              "sofer2015", "todd2016_1", "chou2016", "mani2013", 
              "haimovitz2016_2", "haimovitz2016_1", "craig2014")
options <- filtered |> filter(target_lastauthor_year %in% approved)

#test_that("number of rows in options", {expect_equal(nrow(options),27)})

@@

chosen <- c("birch2007_2", "jara-ettinger2022", "porter_2016_1",
            "hopkins2016", "yeshurun2003", "craig2014", "tarampi2016", 
            "mani2013", "krauss2003", "schechtman2010", "child2018",
            "haimovitz2016_1", "todd2016_1", "haimovitz2016_2", "payne2008", 
            "paxton2012", "dehaene2009", "chou2016", "ngo2019", "gong2019")

rescue_started <- options |> filter(target_lastauthor_year %in% chosen)

```

```{r}
d <- read_csv(here("data", "combined_data.csv")) |> 
  select(target_lastauthor_year, type, on_turk, repeated_measure, N, raw_stat, 
         same_direction, replication_score, closeness, subfield, target_year, stanford_internal, 
         open_data, open_materials, within_between, single_vignette, cost, name_pretty)

source(here("code","helper","parse_stats.R"))
```

```{r}
parsed_d <- d |> 
  mutate(raw_stat=gsub(" ","",raw_stat),
         calc=pmap(list(raw_stat, within_between,N), do_parsing)) |> 
  unnest(cols=c(calc), names_sep="_") |> 
    mutate(
    calc_d_calc=case_when(
      type=="original" ~ abs(calc_d_calc),
      same_direction=="yes" ~ abs(calc_d_calc),
      same_direction=="no" ~ -abs(calc_d_calc),
      T ~ as.numeric(NA)
      ),
    calc_ES=case_when(
      type=="original" ~ abs(calc_ES),
      same_direction=="yes" ~ abs(calc_ES),
      same_direction=="no" ~ -abs(calc_ES),
      T ~ as.numeric(NA)
      ),
    type=factor(type, levels=c("original", "rep1", "rescue", "additional"))
    ) |> 
  rowwise()

num_proj <- parsed_d |> filter(type=="rescue") |> nrow()

cost <- parsed_d |> filter(type=="rescue") |> ungroup() |>  summarize(total=sum(cost, na.rm=T))
```

@@

Imagine you are a graduate student, and you run a replication of a study that you are interested in building upon in your research.
The replication fails.
Perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a $p$-value of .3.
Or perhaps the interaction is numerically in the wrong direction and the main effects look different.
<!--Or maybe the main effect of interest may be null, but you found a significant effect interaction that wasn't hypothesized.--> Whatever the details of the replication failure, you are left with a question: Should you try again and run a re-replication, or should you give up and pick a different study to build upon?

@@

Psychology is not the only discipline where many studies do not replicate.
Large scale replication projects in other disciplines have found 39/97 of studies with positive effects (40%) replicating in cancer biology [@errington2021], 11/18 studies (61%) replicating in economics [@camerer2016], and 31/40 studies (78%) replicating in experimental philosophy [@cova2021].
Across these disciplines, and especially in psychology, large-scale replications indicate a substantial chance of scientists encountering failed replications if they try to replicate published findings.

Replications can fail for many reasons, having to do with the original study or the replication.
<!-- In principle, any difference between the original study and replication could be a root cause of the difference in their outcomes. But in practice, the relationship between original and replication is not symmetric, and many causes of replication failure can be attributed either more to the original study or more to the replication.  --> Starting with the original study, one potential reason for replication failure is that the original result is fragile in some way.
It could be a statistically unlikely result that was achieved through chance or $p$-hacking, or it could be sensitive to the exact conditions and time it was run under ("hidden moderators").
Another possibility is that the original reported effect size might be inflated due to some combination of heterogeneity, $p$-hacking, low power, and publication bias.
A consequence of inflated effect sizes is that replication studies might select a sample size with appropriate statistical power for the reported effect size and then have too small a sample to reliably detect the (true) smaller effect.
Replication studies might also end up underpowered due to unexpected attrition or noise in the replication sample.

Some other reasons for replication failure are primarily due to the replication study.
A replication could observe different results from the original due to methodological differences.
These differences in design could be intentional adaptations, unavoidable changes from lack of access to the original instructions or materials, or changes that were not expected to be critical.
A difference in the details of recruitment (e.g., online versus in-person) might mean that particular details of a study implementation would no longer be appropriate, or the data quality controls and attention checks might be inadequate.
If there are differences in time, place, or subject population between the two studies, corresponding changes to the materials, instructions, or procedure could be needed to provide an adaptation of the experimental paradigm to the new context.
Such adaptation creates yet more challenges in interpretation: too many changes could cause differing results, but so could not enough changes.
In sum, when replications fail, we generally do not know why they failed, although we may speculate.

For the goals of large-scale replication projects, it does not matter why any particular study failed to replicate, because the aim is to estimate the proportion of replication effects in a literature.
In contrast, however, when individual scientists attempt to replicate an effect with the goal of building on it, they do care about why an individual replication failed.
In an uncertain literature, scientists might not want to commit resources to studying extensions of an effect without first checking that they can measure that effect themselves.
For those individual scientists, it is much more important whether the replication failure is due to a fixable issue with the replication or an intrinsic feature of the original result.

Re-replication can serve as a way to test explanations for replication failure.
While it is possible to speculate after the fact about potential causes of a single replication failure, in most circumstances these causes are hard to discern (absent clear evidence of errors).
In contrast, a second replication provides a chance to test explanations by making modifications -- as well as giving another chance for researchers to measure an effect of interest.
But what are the chances that a re-replication succeeds?

@@

Relatively few studies have estimated re-replication success.
One estimate comes from @ebersole2020.
After the original RP:P study reported of a replication rate around 40% [@opensciencecollaboration], @gilbert2016 raised concerns that methodological differences might explain the low replication rate.
In 11 of the RP:P studies, the original authors had raised concerns about protocol fidelity prior to data collection.
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication.
Thus, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts.
This study thus tested whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.
The result was that 2 of the 10 revised protocols found a significant result (but not the 1/11 that originally was successful).
Even when a significant effect was recovered, the effect size was much diminished [@ebersole2020].
<!-- We could frame @ebersole2020 as an attempt to recover the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks. --> The re-replication success rate was thus 2/9 (22%) for studies with non-significant first replications, under favorable, high-resource circumstances.

@@

While @ebersole2020 looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have those kinds of resources to devote to re-replication attempts.
On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure.
Thus, it is an open question what the re-replication rate is for failed replications by early career researchers.
Further, estimating this rate could inform researchers' decisions on what to do next after failed replications.

Here, we investigate re-replication in the context of graduate student replication class projects.
Students in a graduate-level class on experimental methods in psychology were given the option to "rescue" an experiment that a student in a prior year had previously failed to replicate as their class project [first replications reported in @boyce2023].
The focus of the class was best practices in experimental design; thus, students looked at both the original paper and the failed replication to examine potential issues in the replication that they could fix in their rescue project.
In the present paper, we report the results of `r num_proj` re-replication rescue projects.
We find that a minority of projects (5/17, 29%) successfully recovered the original effect.

@@

PSYCH 251 is a graduate-level experimental methods class taught by MCF.
In previous years, students have conducted replication projects [@boyce2023; @hawkins2018].
In Fall 2023, students were offered the option to complete a "rescue" project where they re-replicated one of the unsuccessful replications from a previous year.
Students could also opt to do a normal replication instead.
If they opted for a rescue, they were given the option to participate in a collaborative meta-science project and earn authorship on the current paper based on 1) completing a rescue project, 2) reviewing another student's rescue project for reproducibility, and 3) reviewing and approving the final manuscript.
We report on the result of `r num_proj` rescue projects that opted to be part of the paper and completed data collection.

A spreadsheet of projects, individual project write-ups (both first replications and rescues), individual project data and analyses for rescue projects, and the analytic code for this paper are all available at <https://osf.io/cyk5w/>.

@@

```{r}
rel_size <- parsed_d |> filter(type %in% c("original", "rescue")) |> select(target_lastauthor_year, type, N)  |> 
  pivot_wider(names_from=type, values_from=N) |> mutate(multiplier=rescue/original)

lowest <- quantile(rel_size$multiplier, 0)
low <- quantile(rel_size$multiplier, .25)
mid <- quantile(rel_size$multiplier, .5)
m <- mean(rel_size$multiplier)
high <- quantile(rel_size$multiplier, .75)
highest <- quantile(rel_size$multiplier, 1)

```

```{r}
subj <- parsed_d |> filter(type=="rescue") |> mutate(binary_success=ifelse(replication_score>.5,1,0))

tallied <- subj|> group_by(replication_score) |> tally()

success <- filter(subj, binary_success==1)

d <- read_csv(here("data", "combined_data.csv")) |> filter(type=="rescue")

irr <- cor.test(d$TA_rep_score, d$MCF_rep_score, method="spearman")$estimate

@@

percent_rep <- boyce2023_sample |> mutate(is_rep=ifelse(sub_rep>.5,1,0)) |> summarize(m=mean(is_rep))

overall <- (percent_rep[1])+(1-percent_rep[1])*nrow(success)/nrow(subj)
```

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (<https://osf.io/dfwk2>).
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is less overused but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn.
The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

@@

For the original study and each replication, we recorded the number of participants post-exclusions.
For studies where some extra conditions were dropped in some replications, we counted only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we did not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted online using a crowd-sourced platform or not.

@@

We first discuss the overall rate of replication in the rescue projects, followed by looking at the relative effect sizes of the rescues and correlates of rescue success.
In the last subsection we describe several interesting case studies.

@@

```{r smd, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. In a few cases, the first replication's key effect was in the same direction as the original with a 95% CI that did not overlap 0; however, in these cases the larger pattern of results was not consistent between original and first replication.  "}

#test on just one
colors <- c("Original study"="black", "First replication"="#377EB8", "Rescue"="#E41A1C", "Additional replication"="#984EA3")
shape <- c("Original study"=21,"First replication"=22,"Rescue"=23,"Additional replication"=24)

big <- parsed_d |> filter(type=="original") |> filter(calc_d_calc>1|target_lastauthor_year=="yeshurun2003") |> select(target_lastauthor_year)
for_plotting_d <- parsed_d |> filter(!is.na(calc_d_calc)) |> 
  mutate(point=calc_d_calc, low=calc_d_calc-1.96*calc_d_calc_se, high=calc_d_calc+1.96*calc_d_calc_se) |> 
  mutate(type=case_when(
    type=="original" ~ "Original study",
    type=="rep1" ~ "First replication",
    type=="rescue" ~ "Rescue",
    type=="additional" ~ "Additional replication"
  ),
  type=factor(type, levels=c("Original study", "First replication", "Rescue", "Additional replication")))

smd_scale_small <- ggplot(for_plotting_d |> anti_join(big),aes(x=reorder(name_pretty, point, FUN=max),y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
  scale_shape_manual(values=shape)+
  theme(legend.position = "bottom", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank(),
        legend.title=element_blank(),
        legend.text=element_text(size=10)
        )+
  labs(y="Standardized mean difference", x="")

level_order = c("Yeshurun & Levy 2003",
                "Ngo et al. 2019", 
                 "Krauss & Wang 2003",
                 "Jara-Ettinger et al. 2022")
smd_scale_big <- ggplot(for_plotting_d |> inner_join(big) |> 
                          mutate(name_pretty=factor(name_pretty, levels=level_order)), aes(x=name_pretty,y=point,ymin=low,ymax=high,shape=type, color=type, fill=type, group=desc(type)))+
  geom_errorbar(size=.5,width=0, position=position_dodge(width=.5))+
  geom_point(position=position_dodge(width=.5))+
 coord_flip()+
  scale_size_area()+
  geom_hline(yintercept=0,color="black")+
    scale_color_manual(values=colors, aesthetics=c("color", "fill"))+
    scale_shape_manual(values=shape)+
  theme(legend.position = "none", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x=element_blank()
        )+
  labs(y="Standardized mean difference", x="", )

#smd_scale_big
#smd_scale_small

plot_grid(smd_scale_big, smd_scale_small, rel_heights=c(.4,1), nrow=2)
```

As a complement to subjective replication ratings, we also statistically compared the effect sizes of the rescue, first replication, and original study on one preregistered key measure per study [again following @boyce2023].
Though we did not conduct a systematic review of the literature to identify all possible replications, we also consider the effect sizes obtained in additional replications when we were aware of them (either from other class projects, or external replications in the literature).

@@

```{r}

do_d_rma <- function(df) {
  if(!any(is.na(df$d))){
  r<- rma(yi=df$d, sei=df$d_se, slab=df$type)
  return(tibble(rma_d_est=r$beta[,1], rma_d_se=r$se))
  }
  return(tibble(rma_d_est=NA, rma_d_se=NA))

}

do_p_orig <- function(es, es_se, rep_es, rep_se){
  if(!is.na(es)&!is.na(es_se)&!is.na(rep_es)&!is.na(rep_se)){
    return(Replicate::p_orig(es, es_se**2,rep_es, t2=0.21**2, rep_se**2))
  }
  return(NA)
}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # By default, hide code; set to TRUE to see code
# knitr::opts_chunk$set(fig.pos = 'th') # Places figures at top or here
knitr::opts_chunk$set(
  out.width = "100%", dpi = 300,
  fig.width = 8, fig.width = 8
) # Figure resolution and size
knitr::opts_chunk$set(fig.env = "figure") # Latex figure environment

options(knitr.table.format = "latex") # For kable tables to work without setting format option

knitr::opts_chunk$set(echo = F, warning = F, message = F) # dev = "png", dev.args = list(type = "cairo-png")
library("papaja")
library("bookdown")
library("rticles")
library(here)
library(tidyverse)
library(brms)
library(tidybayes)
library(kableExtra)
library(viridis)
library(cowplot)
library(ggthemes)
library(reshape2)
library(knitr)
library(kableExtra)
library(metafor)
# r_refs("r-references.bib", append=F)
theme_set(theme_bw())
options(knitr.table.format = "pdf")

model_location <- "code/models"
```

```{r}
library(tidyverse)
library(testthat)
all_projects <- read_csv("https://raw.githubusercontent.com/vboyce/251-254-MA/main/data/raw_data.csv")

filtered <- all_projects |>
  mutate(sub_rep = ifelse(replicated_instructor_code == replicated_report_code,
    replicated_report_code,
    adjudicated_replication_code
  )) |>
  filter(sub_rep %in% c(0, .25, .5)) |>
  filter(include != "no") |>
  filter(target_N <= 200) |>
  filter(academic_year %in% c(
    "2015-2016", "2016-2017", "2017-2018", "2018-2019",
    "2019-2020", "2020-2021", "2021-2022", "2022-2023"
  ))
# what we really want is whether there was a github, but that column is private and so not in the public raw data
# but 2015 is the first year where we have github links

# test_that("number of rows in filtered", {expect_equal(nrow(filtered),49)})

# given this sample, we then sought original replicator permission,
# since it would involve sharing their repo and write-up with a new student
# which wouldn't have been anticipated at original classtime

# we heard back from 29 students of the 49 we attempted to contact
# (1 we couldn't contact due to email bouncing)
# 2 responses were negative, rest were positive, for 27 options

approved <- c(
  "krauss2003", "yeshurun2003", "daffner2000", "ngo2019",
  "child2018", "schechtman2010", "payne2008", "paxton2012",
  "hart2018", "lewis2015", "tarampi2016", "jara-ettinger2022",
  "porter_2016_1", "hopkins2016", "birch2007_2", "gong2019",
  "correll2007", "pilditch2019", "daw2011", "dehaene2009",
  "sofer2015", "todd2016_1", "chou2016", "mani2013",
  "haimovitz2016_2", "haimovitz2016_1", "craig2014"
)
options <- filtered |> filter(target_lastauthor_year %in% approved)

# test_that("number of rows in options", {expect_equal(nrow(options),27)})

@@

chosen <- c(
  "birch2007_2", "jara-ettinger2022", "porter_2016_1",
  "hopkins2016", "yeshurun2003", "craig2014", "tarampi2016",
  "mani2013", "krauss2003", "schechtman2010", "child2018",
  "haimovitz2016_1", "todd2016_1", "haimovitz2016_2", "payne2008",
  "paxton2012", "dehaene2009", "chou2016", "ngo2019", "gong2019"
)

rescue_started <- options |> filter(target_lastauthor_year %in% chosen)
```

```{r}
d <- read_csv(here("data", "combined_data.csv")) |>
  select(
    target_lastauthor_year, type, on_turk, repeated_measure, N, raw_stat,
    same_direction, replication_score, closeness, subfield, target_year, stanford_internal,
    open_data, open_materials, within_between, single_vignette, cost, name_pretty
  )

source(here("code", "helper", "parse_stats.R"))
```

```{r}
parsed_d <- d |>
  mutate(
    raw_stat = gsub(" ", "", raw_stat),
    calc = pmap(list(raw_stat, within_between, N), do_parsing)
  ) |>
  unnest(cols = c(calc), names_sep = "_") |>
  mutate(
    calc_d_calc = case_when(
      type == "original" ~ abs(calc_d_calc),
      same_direction == "yes" ~ abs(calc_d_calc),
      same_direction == "no" ~ -abs(calc_d_calc),
      T ~ as.numeric(NA)
    ),
    calc_ES = case_when(
      type == "original" ~ abs(calc_ES),
      same_direction == "yes" ~ abs(calc_ES),
      same_direction == "no" ~ -abs(calc_ES),
      T ~ as.numeric(NA)
    ),
    type = factor(type, levels = c("original", "rep1", "rescue", "additional"))
  ) |>
  rowwise()

num_proj <- parsed_d |>
  filter(type == "rescue") |>
  nrow()

cost <- parsed_d |>
  filter(type == "rescue") |>
  ungroup() |>
  summarize(total = sum(cost, na.rm = T))
```

@@

Imagine you are a graduate student and you run a replication of a study that you are interested in building upon in your research.
The replication fails.
Perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely contains 0, with a $p$-value of .3.
Or perhaps the interaction is numerically in the wrong direction and the main effects look different.
<!--Or maybe the main effect of interest may be null, but you found a significant effect interaction that wasn't hypothesized.--> Whatever the details of the replication failure, you are left with a question: Should you try again and run a re-replication, or should you give up and pick a different study to build upon?

@@

Psychology is not the only discipline where many studies do not replicate.
Large scale replication projects in other disciplines have found 39/97 of studies with positive effects (40%) replicating in cancer biology [@errington2021], 11/18 studies (61%) replicating in economics [@camerer2016], and 31/40 studies (78%) replicating in experimental philosophy [@cova2021].
Across these disciplines, and especially in psychology, large-scale replications indicate that the probability of scientists failing to replicate published findings is substantial.

Replications can fail for many reasons. Some reasons for replication failure have to do with the original study, while others have to do with how the replication was conducted.
<!-- In principle, any difference between the original study and replication could be a root cause of the difference in their outcomes. But in practice, the relationship between original and replication is not symmetric, and many causes of replication failure can be attributed either more to the original study or more to the replication.  --> One potential reason is that the original result is fragile in some way.
It could be a statistically unlikely result that was achieved through chance or $p$-hacking, or it could be sensitive to the exact conditions and time it was run under ("hidden moderators").
Another possibility is that the original reported effect size is inflated due to some combination of heterogeneity, $p$-hacking, low power, and publication bias.
If effect sizes are inflated, replication studies that select a sample size with appropriate statistical power for the reported effect size will have too small a sample to reliably detect the (true) smaller effect.
Replication studies might also end up underpowered due to unexpected attrition or noise in the replication sample.

Other reasons for replication failure are primarily due to the design of the replication study.
A replication could observe different results from the original due to methodological differences.
These differences include intentional adaptations, unavoidable changes from lack of access to the original instructions or materials, or changes that the replicators did not expect to be critical.
A difference in the details of recruitment, such as a change from recruiting participants in person to online, might mean that particular details of a study implementation would no longer be appropriate, or the data quality controls might be inadequate.
If there are differences in time, place, or subject population between the two studies, researchers may need to change the materials, instructions, or procedure to adapt the experimental paradigm to the new context.
Such adaptation creates more challenges in interpretation: too many changes could cause differing results, but so could too few changes.
In sum, when replications fail, we generally do not know why they failed, although we may speculate.

For the goals of large-scale replication projects, it does not matter why any particular study failed to replicate, because the aim is to estimate the proportion of effects that replicate in a literature.
In contrast, when individual scientists attempt to replicate an effect with the goal of building on it, they care why an individual replication failed.
In an uncertain literature, scientists might not want to commit resources to studying extensions of an effect without first checking that they can measure that effect themselves.
For those individual scientists, it is much more important to know whether the replication failure is due to a fixable issue with their replication attempt or an intrinsic feature of the original result.

While it is possible to speculate after the fact about potential causes of a single replication failure, in most circumstances these causes are hard to discern (absent clear evidence of errors).
Re-replication can serve as a way to test explanations for replication failure.
Re-replications can make making modifications to the first replication and give another chance for researchers to measure an effect of interest.
But what are the chances that a re-replication succeeds?

@@

Relatively few studies have estimated re-replication success.
One estimate comes from @ebersole2020.
After the original RP:P study reported of a replication rate around 40% [@opensciencecollaboration], @gilbert2016 raised concerns that methodological differences might explain the low replication rate.
In 11 of the RP:P studies, the original authors had raised concerns about protocol fidelity prior to data collection.
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication.
Thus, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts.
This study thus tested whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.
The result was that 2 of the 10 revised protocols found a significant result (but not the 1/11 that originally was successful).
Even when a significant effect was recovered, the effect size was much diminished [@ebersole2020].
<!-- We could frame @ebersole2020 as an attempt to recover the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks. --> The re-replication success rate was thus 2/9 (22%) for studies with non-significant first replications under favorable, high-resource circumstances.

@@

While @ebersole2020 looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have comparable resources to devote to re-replication attempts.
On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure.
Thus, it is an open question what the re-replication rate is for failed replications by early career researchers.
Further, estimating this rate could inform researchers' decisions on what to do next after failed replications.

Here, we investigate re-replication in the context of graduate student replication class projects.
Students in a graduate-level class on experimental methods in psychology were given the option to "rescue" an experiment that a student in a prior year had previously failed to replicate as their class project [first replications reported in @boyce2023].
The focus of the class was best practices in experimental design; thus, students looked at both the original paper and the failed replication to examine potential issues in the replication that they could fix in their rescue project.
In the present paper, we report the results of `r num_proj` rescue projects.
We find that a minority of projects (5/17, 29%) successfully recovered the original effect.

@@

PSYCH 251 is a graduate-level experimental methods class taught by MCF.
In previous years, students have conducted replication projects [@boyce2023; @hawkins2018].
In Fall 2023, students were offered the option to complete a "rescue" project where they re-replicated one of the unsuccessful replications from a previous year.
Students could also opt to do a normal replication instead.
If they opted for a rescue, they were given the option to participate in a collaborative meta-science project and earn authorship on the current paper based on 1) completing a rescue project, 2) reviewing another student's rescue project for reproducibility, and 3) reviewing and approving the final manuscript.
We report on the result of the `r num_proj` rescue projects by students who completed data collection and opted to be a part of the paper.

A spreadsheet of projects, individual project write-ups (both first replications and rescues), individual project data and analyses for rescue projects, and the analytic code for this paper are all available at <https://osf.io/cyk5w>.

@@

```{r}
rel_size <- parsed_d |>
  filter(type %in% c("original", "rescue")) |>
  select(target_lastauthor_year, type, N) |>
  pivot_wider(names_from = type, values_from = N) |>
  mutate(multiplier = rescue / original)

lowest <- quantile(rel_size$multiplier, 0)
low <- quantile(rel_size$multiplier, .25)
mid <- quantile(rel_size$multiplier, .5)
m <- mean(rel_size$multiplier)
high <- quantile(rel_size$multiplier, .75)
highest <- quantile(rel_size$multiplier, 1)
```

```{r}
subj <- parsed_d |>
  filter(type == "rescue") |>
  mutate(binary_success = ifelse(replication_score > .5, 1, 0))

tallied <- subj |>
  group_by(replication_score) |>
  tally()

success <- filter(subj, binary_success == 1)

d <- read_csv(here("data", "combined_data.csv")) |> filter(type == "rescue")

irr <- cor.test(d$TA_rep_score, d$MCF_rep_score, method = "spearman")$estimate

@@

percent_rep <- boyce2023_sample |>
  mutate(is_rep = ifelse(sub_rep > .5, 1, 0)) |>
  summarize(m = mean(is_rep))

overall <- (percent_rep[1]) + (1 - percent_rep[1]) * nrow(success) / nrow(subj)
```

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (<https://osf.io/dfwk2>).
In general, we encouraged students to add manipulation checks as appropriate and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of @paxton2012 switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is used less but attempts to measure the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn.
The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

@@

For the original study and each replication, we recorded the number of participants after exclusions.
For studies where extra conditions were dropped in some replications, we counted only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we did not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted online using a crowd-sourced platform or not.

@@

We first discuss the overall rate of replication in the rescue projects, then we look at the relative effect sizes of the rescues and correlates of rescue success.
In the last subsection we describe several interesting case studies.

@@

```{r smd, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of some studies, studies with large effects are shown in a separate panel. In a few cases, the first replication's key effect was in the same direction as the original with a 95% CI that did not contain 0. However, in these cases the larger pattern of results was not consistent between original and first replication."}
# test on just one
colors <- c("Original study" = "black", "First replication" = "#377EB8", "Rescue" = "#E41A1C", "Additional replication" = "#984EA3")
shape <- c("Original study" = 21, "First replication" = 22, "Rescue" = 23, "Additional replication" = 24)

big <- parsed_d |>
  filter(type == "original") |>
  filter(calc_d_calc > 1 | target_lastauthor_year == "yeshurun2003") |>
  select(target_lastauthor_year)
for_plotting_d <- parsed_d |>
  filter(!is.na(calc_d_calc)) |>
  mutate(point = calc_d_calc, low = calc_d_calc - 1.96 * calc_d_calc_se, high = calc_d_calc + 1.96 * calc_d_calc_se) |>
  mutate(
    type = case_when(
      type == "original" ~ "Original study",
      type == "rep1" ~ "First replication",
      type == "rescue" ~ "Rescue",
      type == "additional" ~ "Additional replication"
    ),
    type = factor(type, levels = c("Original study", "First replication", "Rescue", "Additional replication"))
  )

smd_scale_small <- ggplot(for_plotting_d |> anti_join(big), aes(x = reorder(name_pretty, point, FUN = max), y = point, ymin = low, ymax = high, shape = type, color = type, fill = type, group = desc(type))) +
  geom_errorbar(size = .5, width = 0, position = position_dodge(width = .5)) +
  geom_point(position = position_dodge(width = .5)) +
  coord_flip() +
  scale_size_area() +
  geom_hline(yintercept = 0, color = "black") +
  scale_color_manual(values = colors, aesthetics = c("color", "fill")) +
  scale_shape_manual(values = shape) +
  theme(
    legend.position = "bottom",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 10)
  ) +
  labs(y = "Standardized mean difference", x = "")

level_order <- c(
  "Yeshurun & Levy 2003",
  "Ngo et al. 2019",
  "Krauss & Wang 2003",
  "Jara-Ettinger et al. 2022"
)
smd_scale_big <- ggplot(for_plotting_d |> inner_join(big) |>
  mutate(name_pretty = factor(name_pretty, levels = level_order)), aes(x = name_pretty, y = point, ymin = low, ymax = high, shape = type, color = type, fill = type, group = desc(type))) +
  geom_errorbar(size = .5, width = 0, position = position_dodge(width = .5)) +
  geom_point(position = position_dodge(width = .5)) +
  coord_flip() +
  scale_size_area() +
  geom_hline(yintercept = 0, color = "black") +
  scale_color_manual(values = colors, aesthetics = c("color", "fill")) +
  scale_shape_manual(values = shape) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) +
  labs(y = "Standardized mean difference", x = "", )

# smd_scale_big
# smd_scale_small

plot_grid(smd_scale_big, smd_scale_small, rel_heights = c(.4, 1), nrow = 2)
```

As a complement to subjective replication ratings, we also statistically compared the effect sizes of the rescue, first replication, and original study on one preregistered key measure per study [again following @boyce2023].
Though we did not conduct a systematic review of the literature to identify all possible replications, we also considered the effect sizes obtained in additional replications when we were aware of them (either from other class projects, or external replications in the literature).

@@

```{r}
do_d_rma <- function(df) {
  if (!any(is.na(df$d))) {
    r <- rma(yi = df$d, sei = df$d_se, slab = df$type)
    return(tibble(rma_d_est = r$beta[, 1], rma_d_se = r$se))
  }
  return(tibble(rma_d_est = NA, rma_d_se = NA))
}

do_p_orig <- function(es, es_se, rep_es, rep_se) {
  if (!is.na(es) & !is.na(es_se) & !is.na(rep_es) & !is.na(rep_se)) {
    return(Replicate::p_orig(es, es_se**2, rep_es, t2 = 0.21**2, rep_se**2))
  }
  return(NA)
}

orig <- parsed_d |>
  filter(type == "original") |>
  select(target_lastauthor_year, d = calc_d_calc, d_se = calc_d_calc_se)
